### 2012-10-30 4.tex
Graphs and tables

\chapter{Structural and accidental gaps} 
\label{gaps}

@@

@@

Unlike prior studies, ``neo-classical compounds'', words which appear to consist of a Latinate prefix and a bound stem (e.g., \emph{inspect}, \emph{excrete}) are also excluded from this wordlist. This is done in response to converging evidence that English speakers regard such words as complex. This assumption has long been made to simplify phonological \citep{SPE}, allomorphic \citep[11f.]{Aronoff1976}, and syntactic \citep{Harley2009} generalizations. Results from lexical decision experiments also suggest that neoclassical compounds are complex. \citet{Taft1986} and \citet{Taft1975,Taft1976} find that nonce words like \emph{*re-sert}, which appear to be composed of a prefix and a bound stem, take longer to reject that non-words which lack apparent morphological structure such as *\emph{refant}). Bound stems also show frequency effects which are independent of their whole word frequency \citep{Taft1979,Taft2006}. Finally, \citet{Emmorey1989} and \citet{Forster2000} report facilitative priming between pairs like \emph{permit}-\emph{submit}, an effect suggestive of a shared morphological identity.

@@

In summary, there is no statistical support for any of the three static constraints proposed by \citeauthor{Pierrehumbert1994}. However, it is premature to conclude that no static constraint exists.

%It is possible that an explicit model of ``constraint discovery'', could identify reliable static constraints,

@@

The allomorphs of the regular past (e.g., \emph{nap}[t] $\sim$ \emph{nab}[d]) and of the regular noun plural (e.g., \emph{lap}[s] $\sim$ \emph{lab}[z]) illustrate the well-known process of \textsc{Obstruent Voice Assimilation} (e.g., \emph{SPE}:178) in which the voicing specification of the first of two adjacent obstruents perserveres.

\begin{example}[\textsc{Obstruent Voice Assimilation}]
$\begin{bmatrix} -\alpha~\textsc{Voiced} \\ -\textsc{Sonorant} \end{bmatrix}~\goesto{}~\begin{bmatrix} \alpha~\textsc{Voiced} \\ -\textsc{Sonorant} \end{bmatrix}~\big /~\begin{bmatrix} \alpha~\textsc{Voiced} \\ -\textsc{Sonorant} \end{bmatrix}~\gap{}$
\end{example}

Though it is not mentioned by \citet{Pierrehumbert1994} in her study of medial clusters, the vast majority of medial obstruent clusters are either uniformly voiced or voiceless, as in \emph{hu}[z.b]\emph{and} or \emph{rha}[p.s]\emph{osdy}. ``Disagreeing'' clusters, like those in \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos} are rare

@@

%``Neo-classical compounds'', words that appear to consist of a Latinate prefix and a bound stem, like \emph{inspect} or \emph{excrete}, have an ambiguous and long-debated morphological status. \citeauthor{Pierrehumbert1994} attempts to distinguish between words like \emph{complete}, \emph{extreme}, \emph{inspect}, \emph{obtuse}, which she assumes to be simplex, and \emph{excrete} and \emph{transparent}, which she considers complex. However, CELEX codes these words as complex and they are thus excluded. This is consistent with converging evidence that speakers decompose these Latinate forms. The baroque treatment of Latinate forms in \emph{The sound pattern of English} (\citealp{SPE}, henceforth \emph{SPE}) assumes a prefix/bound stem decomposition to simplify various morphophonological generalizations.  In a similar vein, \citet[11--13]{Aronoff1976} observes that Latinate forms which share the same bound stem also share irregular allomorphs of that stem under derivation, which is presented as evidence of prefix/bound stem decomposition.
%\begin{example}[Bound stem-specific allomorphy]
%a. & {adhere}   & {adhesion}   \\
%   & {cohere}   & {cohesion}   \\
%b. & {conceive} & {conception} \\
%   & {perceive} & {perception} \\
%There is also a syntactic interaction between Latinate prefixes on verbs and the makeup of the verbal complement. \citet{Harley2009} observes that Latinate verbs do not generally participate in ditransitive, verb participle, or adjectival resultative constructions, all acceptable with similar Anglo-Saxon verbs \citep[see also][]{Gropen1989,Coppock2008}. \citeauthor{Harley2009} proposes that the Latinate prefix is introduced in a low small clause, the same position as the theme, participle, and result adjectival in the Anglo-Saxon examples.
%\begin{example}[Latinate verbs and small clauses] 
%\label{harley}
%\begin{tabular}{ll@{}ll@{}l}
%a. & * & {exhibit him the painting} & ~ & {show him the painting} \\
%b. & * & {imbibe himself stupid}    & ~ & {drink himself stupid}  \\
%c. & * & {exhibit it off}           & ~ & {show it off}           \\

@@

%The [t \alt{} d] allomorphs of the regular past (e.g. \emph{nap}[t] $\sim$ \emph{nab}[d]), and the [s \alt{} z] allomorphs of the regular noun plural (e.g., \emph{lap}[s] $\sim$ \emph{lab}[z]), Saxon genitive, and third person singular active indicative verb agreement, exemplify \textsc{Obstruent Voice Assimilation} (\emph{SPE}:178).\footnote{Underlying /-d, -z/ are argued for by \citet[282]{Hockett1958}, \citet[210]{SPE}, \citet{Basboll1972}, \citet{Shibatani1972}, \citet{Anderson1973a}, \citet[102]{Pinker1988}, and \citet[284f.]{Bakovic2005b}; alternatives are suggested by \citet[210f.]{LANGUAGE}, \citet[426]{Nida1948}, \citet{Luelsdorff1969}, \citet{Lightner1970}, \citet{Hoard1971}, \citet{Miner1975}, \citet{Zwicky1975}, \citet{Kiparsky1985}, and \citet[135]{Borowsky1986}.}

@@

%$\begin{bmatrix} -\textsc{Son} \end{bmatrix}~\goesto~\begin{bmatrix} =\textsc{Voi} \end{bmatrix}~/~\gap~\begin{bmatrix} =\textsc{Voi} \\ -\textsc{Son} \end{bmatrix}$

@@

%\label{npa}
%In addition to [n $\sim$ ŋ] allophony, \textsc{Nasal Place Assimilation} (\emph{SPE}:85) is responsible for the allomorphs of the \emph{im-}/\emph{in-} prefix, determined by the place of the following consonant.
%\begin{example}[\emph{im-}/\emph{in-} allomorphy] 
%\label{nparule}
%\begin{tabular}{l l l l l l l}
%a. & {polite}   & {i}[m.p]{olite}   & & {balance} & {i}[m.b]{alance} \\
%b. & {tangible} & {i}[n.t]{angible} & & {decent}  & {i}[n.d]{ecent}  \\
%% mention Jensen here
%\begin{example}[\textsc{Nasal Place Assimilation}]
%%                          & \txt{\textsc{Place}}\ar@{-}[d]\ar@{--}[dl] \\
%%\txt{[$+$\textsc{Nasal}]} & \txt{[$-$\textsc{Sonorant}]}              \\
%$\begin{bmatrix} $+$\textsc{Nas} \end{bmatrix}~\goesto~\begin{bmatrix} =\textsc{Lab} \\ =\textsc{Cor} \\ =\textsc{Dor} \end{bmatrix}~/~\gap{}~\begin{bmatrix} =\textsc{Lab} \\ =\textsc{Cor} \\ =\textsc{Dor} \\ -\textsc{Son} \end{bmatrix}$

>>>>

\chapter{Structural and accidental phonotatic gaps}
\label{gaps}

@@

\begin{quote}
It would be a result of the greatest interest if it were to turn out that every intramorphemic regularity was necessarily reflected in an alternation, and thus in a rule; but this position cannot seriously be maintained. \citep[283]{Anderson1974}
\end{quote}

@@

Unlike prior studies, ``neo-classical compounds'', words which appear to consist of a Latinate prefix and a bound stem (e.g., \emph{inspect}, \emph{excrete}) are also excluded from this wordlist. This is done in response to converging evidence that English speakers regard such words as complex.

The baroque treatment of Latinate forms in \emph{The sound pattern of English} (\citealp{SPE}, henceforth \emph{SPE}) assumes a prefix/bound stem decomposition to simplify various morphophonological generalizations.

In a similar vein, \citet[11f.]{Aronoff1976} observes that Latinate forms which share the same bound stem also share irregular allomorphs of that stem under derivation, which is presented as evidence of prefix/bound stem decomposition.

\begin{example}[Bound stem-specific allomorphy]
\begin{tabular}{l l l l l l}
a. & {adhere}   & {adhesion}   \\
   & {cohere}   & {cohesion}   \\
b. & {conceive} & {conception} \\
   & {perceive} & {perception} \\
\end{tabular}
\end{example}

This assumption has long been made to simplify phonological \citep{SPE}, allomorphic \citep[11f.]{Aronoff1976}, and syntactic \citep{Harley2009} generalizations.

Results from lexical decision experiments also suggest that neoclassical compounds are complex. \citet{Taft1986} and \citet{Taft1975,Taft1976} find that nonce words like \emph{*re-sert}, which appear to be composed of a prefix and a bound stem, take longer to reject that non-words which lack apparent morphological structure such as *\emph{refant}). Bound stems also show frequency effects which are independent of their whole word frequency \citep{Taft1979,Taft2006}. Finally, \citet{Emmorey1989} and \citet{Forster2000} report facilitative priming between pairs like \emph{permit}-\emph{submit}, an effect suggestive of a shared morphological identity.

There is also a syntactic interaction between Latinate prefixes on verbs and the makeup of the verbal complement. \citet{Harley2009} observes that Latinate verbs do not generally participate in ditransitive, verb participle, or adjectival resultative constructions, all acceptable with similar Anglo-Saxon verbs \citep[see also][]{Gropen1989,Coppock2008}. \citeauthor{Harley2009} proposes that the Latinate prefix is introduced in a low small clause, the same position as the theme, participle, and result adjectival in the Anglo-Saxon examples.

\begin{example}[Latinate verbs and small clauses] 
\label{harley}
\begin{tabular}{l l l l@{} l}
a. & {show him the painting} & \alt{} & * & {exhibit him the painting} \\
b. & {drink himself stupid}  & \alt{} & * & {imbibe himself stupid}    \\
c. & {show it off}           & \alt{} & * & {exhibit it off}           \\
\end{tabular}
\end{example}

@@

There is no statistical support for any of the three static constraints proposed by \citeauthor{Pierrehumbert1994}. 
%However, it is premature to conclude that no static constraint exists. It is possible that an explicit model of ``constraint discovery'', could identify reliable static constraints,

@@

\paragraph{Obstruent voice assimilation}

Voice assimilation alternations are evidenced by the non-syllabic allomorphs of the regular past (e.g., \emph{nap}[t] $\sim$ \emph{nab}[d]) and noun plural (e.g., \emph{lap}[s] $\sim$ \emph{lab}[z]), which take the voicing specification of a preceding obstruent.\footnote{Underlying /-d, -z/ are assumed here (e.g., \citealt{Anderson1973a}, \citealt[284f.]{Bakovic2005b}, \citealt{Basboll1972}, \citealt[210]{SPE}, \citealt[282]{Hockett1958}, \citealt[102]{Pinker1988}, \citealt{Shibatani1972}); alternative analyses are put forth by \citet[210f.]{LANGUAGE}, \citet[135]{Borowsky1986}, \citet{Hoard1971}, \citet{Kiparsky1985}, \citet{Lightner1970}, \citet{Luelsdorff1969}, \citet{Miner1975}, \citet[426]{Nida1948}, and \citet{Zwicky1975}.}

\begin{example}[\textsc{Obstruent Voice Assimilation}]
\label{ovarule}
$\begin{bmatrix} -\textsc{Son} \end{bmatrix}~\goesto~\begin{bmatrix} =\textsc{Voi} \end{bmatrix}~/~\gap~\begin{bmatrix} =\textsc{Voi} \\ -\textsc{Son} \end{bmatrix}$
\end{example}

\noindent
\citet{Pierrehumbert1994} does not discuss a constraint against adjacent obstruents disagreeing in voice. However, the vast majority of medial obstruents clusters in simplex wors are either uniformly voiced, as in \emph{hu}[z.b]\emph{and}, or uniformly voiceless, as in or \emph{rha}[p.s]\emph{osdy}. Hetero-voiced clusters, like those in \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos} are rare in comparison.

\begin{table}
\centering
\begin{tabular}{l rrrr}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 35       & 329        & 10\%       & \multirow{2}{*}{0.002} \\
violating  & 11       & 305        &  3\%       \\
\bottomrule
\end{tabular}
\caption{FIXME}
\label{ovatab}
\end{table}

\paragraph{Nasal place assimilation}
\label{npa}

\textsc{Nasal Place Assimilation} (\emph{SPE}:85) accounts for [n $\sim$ ŋ] allophony as well as

is responsible for the allomorphs of the \emph{im-}/\emph{in-} prefix, determined by the place of the following consonant (though there are additional complexities; see \citealt{Borowsky1986}:?, \citealt{Halle1985a}, \citealt{Jensen2000}).

\begin{example}[\emph{im-}/\emph{in-} allomorphy]
\label{nparule}
\begin{tabular}{l l l l l l l}
a. & {polite}   & {i}[m.p]{olite}   & & {balance} & {i}[m.b]{alance} \\
b. & {tangible} & {i}[n.t]{angible} & & {decent}  & {i}[n.d]{ecent}  \\
\end{tabular}
\end{example}

\begin{example}[\textsc{Nasal Place Assimilation}]
$\begin{bmatrix} $+$\textsc{Nas} \end{bmatrix}~\goesto~\begin{bmatrix} =\textsc{Lab} \\ =\textsc{Cor} \\ =\textsc{Dor} \end{bmatrix}~/~\gap{}~\begin{bmatrix} =\textsc{Lab} \\ =\textsc{Cor} \\ =\textsc{Dor} \\ -\textsc{Son} \end{bmatrix}$
\end{example}

\begin{table}
\centering
\begin{tabular}{l rrrr}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 31       & 2          & 94\%       & \multirow{2}{*}{3.37\e{-07}} \\
violating  & 11       & 22         & 33\%       \\
\bottomrule
\end{tabular}
\caption{FIXME}
\label{npatab}
\end{table}

\paragraph{Degemination}

The final alternation found in English medial clusters is the simplification of geminates characteristic of ``level I'' morphology. Verb roots ending in /d/ which select the irregular /-t/ past tense suffix exhibit degemination (e.g., \emph{bend}/\emph{ben}[t], \emph{build}/\emph{buil}[t]).

A final alternation targeting English syllable contact clusters is the simplification of geminates certain affixes.

The deadjectival suffix \emph{-ly} does not result in a geminate \emph{l} when it attaches to

krarely produces a geminate when it attaches to /l/-final roots (e.g., \emph{norma}[l]\emph{y}, cf. \emph{calm}[l]\emph{y}).

More complicated cases of \textsc{Degemination} can be found in Latinate prefix allomorphy (\emph{SPE}:148; \citealp[102]{Borowsky1986}).

\begin{example}[\textsc{Degemination}]
$\begin{bmatrix} =\textsc{Lab} \\ =\textsc{Cor} \\ =\textsc{Dor} \\ =\textsc{Son} \\ \ldots{} \end{bmatrix}~\goesto~\emptyset~\big /~\gap~\begin{bmatrix} =\textsc{Lab} \\ =\textsc{Cor} \\ =\textsc{Dor} \\ =\textsc{Son} \\ \ldots{} \end{bmatrix}$
\end{example}

\begin{table}
\centering
\begin{tabular}{l rrrr}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 173      & 643        & 21\%       & \multirow{2}{*}{1.2\e{-10}} \\
violating  & 0        & 104        & 0\%        \\
\bottomrule
\end{tabular}
\caption{FIXME}
\label{degemtab}
\end{table}

@@

%``Neo-classical compounds'', words that appear to consist of a Latinate prefix and a bound stem, like \emph{inspect} or \emph{excrete}, have an ambiguous and long-debated morphological status. \citeauthor{Pierrehumbert1994} attempts to distinguish between words like \emph{complete}, \emph{extreme}, \emph{inspect}, \emph{obtuse}, which she assumes to be simplex, and \emph{excrete} and \emph{transparent}, which she considers complex. However, CELEX codes these words as complex and they are thus excluded. This is consistent with converging evidence that speakers decompose these Latinate forms.

@@

@@

@@
